Insurance choice, decision assistance, and insurer steering

### 2021-03-25 finals/lunch-and-learn/lunch-learn-202103.Rmd
update figures and presentation

---
title: "Insurance choice, decision assistance, and insurer steering"
subtitle: "Ian McCarthy & Evan Saltzman"
author: Emory University, Lunch and Learn
date: March 30, 2021 #"`r format(Sys.time(), '%d %B %Y')`"
header-includes: 
  - \usepackage{graphicx}
  - \usepackage{amsmath}
output:
#  html_document: default (toggle on for "simplified" view)
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, custom.css] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      beforeInit: "macros.js"
#knit: pagedown::chrome_print
---

@@

  - "More than half of all uninsured (56%) say they are likely to go to the marketplace to find a health plan in the future. But there are knowledge gaps."
  - "Six in 10 are confused about or have not heard of the tax credit, and more than 7 in 10 want one-on-one help understanding their insurance options"

@@

- 12.7 million enrollees on exchanges in 2016
- Down to 11.4 million in 2020
- Over 27 million remain uninsured

@@

- Large number of people still uninsured
- Many eligible for some type of subsidy on exchanges
]

@@

1. Does some form of decision assistance "help" in plan selection?
2. Does the type of assistance matter (private versus public)?
3. Do insurers use brokers/agents to steer people to different plans?
4. What are the welfare effects of any such steering?

@@

@@

>>>>

---
title: "Insurance choice, decision assistance, and insurer steering"
subtitle: "Ian McCarthy & Evan Saltzman"
author: Emory University, Lunch and Learn
date: March 30, 2021 #"`r format(Sys.time(), '%d %B %Y')`"
header-includes: 
  - \usepackage{graphicx}
  - \usepackage{amsmath}
output:
#  html_document: default (toggle on for "simplified" view)
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, custom.css] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      beforeInit: ["macros.js","https://platform.twitter.com/widgets.js"]
#knit: pagedown::chrome_print
---

@@

- "More than half of all uninsured (56%) say they are likely to go to the marketplace to find a health plan in the future. But there are knowledge gaps."

- "Six in 10 are confused about or have not heard of the tax credit, and more than 7 in 10 want one-on-one help understanding their insurance options"

@@

- 12.7 million enrollees on exchanges in 2016

- Down to 11.4 million in 2020

- Over 27 million remain uninsured

@@

- Large number of people still uninsured

- Many eligible for some type of subsidy on exchanges
]

@@

1. Does some form of decision assistance "help" in plan selection?

2. Do insurers use brokers/agents to steer people to different plans?

3. What are the welfare effects of any such steering?

@@

---
# Sources

Main enrollment come from Covered California under FOIA. This includes:
- Individual demographics and income
- Plan purchases
- Information on type of assistance (if any)

Supplemental data include:
- ACS and SIPP for predicting uninsured
- ACA program data (contribution percentages, rating areas, etc.)

---
# Final data

All analysis done at the household level...

- **Initial data:** 16.9 million households from 2014-2019
- Remove households with missing or inconsistent data
  - 14.6 million remaining
- Remove households predicted to be insured elsewhere
  - 8.3 million remaining
  - 4.5 million of which are new enrollees

---
# Enrollees over time

.center[
  ![:scale 50%](../../figures/enrollee_count.png)
]

---
# Insurers

.pull-left[
  ![](../../figures/insurer_stack_any.png)
]

.pull-right[
  ![](../../figures/insurer_stack_all.png)
]

---
# Metal tier

.pull-left[
  ![](../../figures/metal_stack_any.png)
]

.pull-right[
  ![](../../figures/metal_stack_all.png)

]

@@

---
# What is a dominated choice?

.center[
  ![](https://twitter.com/DaveEvansPhD/status/1234506064470716416?s=20)
]
